---
title: "Functions and help (tasks)"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = param, eval= param, include=T)
```



# Round and signif {#signif}

- Use the `signif` function to print `pi` with 4 significant digits (4 decimal places)

```{r}
signif(pi, 4)
```

- (*) What is the value of `round(pi, -1)` and why?

```{r}
round(pi,-1)
```

- (*) The function `floor()` rounds down a number, whilst `trunc()` truncates a number down to zero. For what `x` is `trunc(x)` different from `floor(x)`?

```{r}
x <- c(-2, -1.5, -1, -0.5, 0, 0.5, 1 )
trunc(x)
floor(x)
```


## `barplot` options {#help}

Have a look at the `women`  data. It reports average heights and weights for groups of North-American women aged 30-39.

- Open the help file for `women` to understand the data.

```{r eval=FALSE}
?women
```

- The `barplot` function can be used for barplots. Make a barplot of `weight`.

```{r}
barplot(women$weight)
```

-	What argument of `barplot` can be used to add names to the bars? HINT: `names.arg`. Add `height` as a names under the bars.

```{r}
barplot(women$weight, names.arg=women$height)
```

- Basic colors are represented in R simply by their names, e.g. `"yellow"`. Note the quotation marks. Make the bars in the plot purple.

```{r}
barplot(women$weight, names.arg=women$height, col="purple")
```

- Add labels to the plot and a title

```{r}
barplot(women$weight, names.arg=women$height, xlab='height', ylab='weight', main='Average weight vs. height') 
```


## Spearman correlation

In the `survey` data:

```{r echo = TRUE, eval = FALSE }
survey <- read.delim( "./data/survey.txt" )
```

```{r echo = FALSE, eval = TRUE }
survey <- read.delim( "../../data/survey.txt" )
```

-	The function for correlation in R is `cor()`. Calculate correlation between average `span1` and `span2`

```{r}
cor(survey$span1, survey$span1)
```

- Is this a Pearson or Spearman correlation? Calculate the Spearman correlation.

```{r}
cor(survey$span1, survey$span1, method='spearman')
```

# (*) `table` options

`NA` in R stands for "not available", that is, it indicates missing values

- Open the help file for `table`.

```{r eval=FALSE}
?table
```

- What is the default of the `useNA` argument?

```{r}
"no"
```

- Use `table()` to the `clap` and `m.i.`columns in `survey` with all three options for `useNA`

```{r}
table(survey$m.i, useNA = 'no')
table(survey$m.i, useNA = 'ifany')
table(survey$m.i, useNA = 'always')
table(survey$clap, useNA = 'no')
table(survey$clap, useNA = 'ifany')
table(survey$clap, useNA = 'always')
```

- Explain the role of the `useNA` argument in your own words.
