---
title: "Matrices (tasks)"
output: html_document
---

```{r setup, include=FALSE,eval=TRUE}
source( "../../setup/R_course_setup.R" );
```

# Matrices (tasks)

## Tasks

### Creation

Create a matrix with 7 rows and 4 columns of the sequence of numbers from 1 until 28, filling out each column at a time.

```{r}
mat <- matrix(1:28, nrow = 7, ncol = 4)
```

### Setting/getting names of rows/columns

Assign to the matrix row names corresponding to weekdays: "Monday", "Tuesday", etc.

```{r}
rownames(mat) <- c("Monday", "Tuesday", "Wednesday",
                   "Thursday", "Friday", "Saturday", "Sunday")
```

Assign to the matrix column names corresponding to week numbers sequentially starting from 3.

```{r}
colnames(mat) <- 3:6
```

### Getting matrix elements

Select from this matrix only the rows corresponding to weekend days, and save the result as an R object. Check that the result is correct.

```{r}
weekend <- mat[c("Saturday", "Sunday"), ]
weekend
```

### Useful matrix functions

The matrix below contains the daily number of loafs of bread sold by a small bakery, for 28 consecutive days.


```{r}
nloafs <- c(30, 24, 24, 27, 33, 45, 60, 32, 25, 22, 26, 34, 43, 61,
            28, 25, 23, 28, 32, 46, 58, 31, 23, 25, 28, 34, 42, 63)
mat <- matrix(nloafs, nrow = 7, ncol = 4)
rownames(mat) <- c("Monday", "Tuesday", "Wednesday",
                   "Thursday", "Friday", "Saturday", "Sunday")
colnames(mat) <- 3:6
```

Compute the total number of loafs sold for each of the weeks.

```{r}
colSums(mat)
```

Compute the total number of loafs sold for each weekday.

```{r}
rowSums(mat)
```

(*) Compute the mean number fo loafs sold for each weekday, across the 4 weeks.

```{r}
rowSums(mat)/4
rowMeans(mat) # gives the same result
```


### Matrix of correlation coefficients {#corMatrix}

#### Extra tasks

The goal of this exercise is to calculate a matrix of column-column Pearson correlation coefficients for a number of data frame columns. The `cor` function does not accept a `data.frame` argument and it works only with numerical columns. Therefore, first the numerical columns of the data frame need to be extracted and converted to a matrix.

- Read the pulse data frame to `pulse` variable; check its class.
```{r}
pulse <- read.table( "../../data/pulse.txt" )
```
```{r result="as.is", echo=FALSE, comment=""}
catReadTable( "pulse.txt", "pulse" )
```

```{r}
class( pulse )
```

- Calculate correlation of columns `height` and `weight`.
```{r}
cor( pulse$height, pulse$weight )
```

- Understand the error of `cor( pulse )`.
```{r error=TRUE}
cor( pulse )
```

- Select only numerical columns of `pulse`. Then use `as.matrix` on the selected columns. Store the result in `m`. Confirm that `m` is a matrix. Check `m` dimensions and `str( m )`.
```{r}
m <- as.matrix( pulse[ c( "height", "weight", "age", "pulse1", "pulse2" ) ] )
class( m )
dim( m )
str( m )
```

- Try `cor( m )` and understand the result. Read the `cor` documentation for further ideas related to `NA`. Assign the final result to `corM`.
```{r}
cor( m )
cor( m, use = 'p' )
corM <- cor( m, use = 'p' )
```

- Read documentation for `round` and round the `corM` matrix to 2 most significant digits. (*)
```{r}
round( corM, 2 )
```
